---
title: "Practical Machine Learning Course Project"
author: "Priscilla Kurtz"
date: "Thursday, December 18, 2014"
output: html_document
---

This is the course project for the Practical Machine Learning from the Data Science Specialization. The goal of your project is to predict the manner in which they did the exercise.

# Prediction Study Design

First we need to load the data and do some cleaning.

```{r}
library(caret)

testing <- read.csv(file = "pml-testing.csv", header = TRUE, sep = ",", na.strings=c("",".","NA"))
training <- read.csv(file = "pml-training.csv",header = TRUE, sep = ",", na.strings=c("",".","NA"))

featurePlot(x = training, y = training$X, plot="pairs")
```

For cleaning, I chose to use the ```preProcess``` function to standardize the data:
```{r}
testing_p <- preProcessr po
```

You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 